---
title: "neualNetSales"
author: "Dr. J"
date: "7/21/2019"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: united
---

# Introduction
This is an application of a neural network designed to identify characterists of car buyers.  The purose is to direct advertising

**ASSIGNMENT:**    
  1. Find the 'best' sales target price for predicting sales -  
  2. Determing values of age, gender, income level or commute distance *i.e.* miles/wk that predict best sales  
  3. Determing the appropriate neural net - how many hidden layers, what variables to leave out?  
  4. Does it help to use 80/20 (or any other) split for training/test sets?  
  5. Recommend an advetising approach based on your findings  
  
## first set up the environment

<!-- Assumes the .csv file 'cars.csv' is in the working directory !--> 

```{r setup, include=FALSE}
set.seed(500)
knitr::opts_chunk$set(echo = TRUE)
library(neuralnet)
cars = read.csv('cars.csv', header=TRUE)
```

## table ..  description of the data
The data set contains information about car buyers

Variable       |  Description
---------------|----------------------------------------
age            | age of the buyer
gender         | sex of the buyer
miles          | average number of miles driven per day
debt           | current debt of the buyer
income         | buyer montly income
sales          | amount spent on a used car


## set up the normalized data:  

```{r}
#normalize age, miles, debt, income, and sales
carsNormalized = cars
carsNormalized$age = (cars$age-mean(cars$age))/sd(cars$age)
carsNormalized$miles = (cars$miles-mean(cars$miles))/sd(cars$miles)
carsNormalized$debt = (cars$debt - mean(cars$debt))/sd(cars$debt)
carsNormalized$income = (cars$income-mean(cars$income))/sd(cars$income)
carsNormalized$sales = (cars$sales-mean(cars$sales))/sd(cars$sales)

```

## print the first 6 entries of the cars data set

=====  
set the target price for sales  
=====  

```{r}

salesTargetPrice = 8500

summary(cars$sales)
hist(cars$sales, xlab='price', main='Used Car Sales Prices')

# set up 'yes / no' based on sales price salesTargetPrice
cars$salesTgt = ifelse(cars$sales>salesTargetPrice,1,0)
carsNormalized$salesTgt = ifelse(cars$sales>salesTargetPrice,1,0)

head(cars)
```

# Set up the Training and Test data  

=====    
set the training fraction  
=====


```{r}
##set up the training and test data:

trainingFraction = 0.75

lengthData = length(cars$age) #  how many entries in the data set
nTrain = 0
nTest=0
nTrain = round(trainingFraction*lengthData,0)
nTest = lengthData-nTrain

# shuffle the data, take the first nTrain as the training set, the 
# remainder as the test set
shuffle = sample(1:lengthData) 

testDataSet = vector("numeric", length=nTest)
trainDataset= vector("numeric", length=nTrain)
trainIndex = shuffle[1:nTrain]
trainDataSet = cars[shuffle[1:nTrain],]
normalizedTrainDataSet= carsNormalized[shuffle[1:nTrain],]

testIndex = vector("numeric")
testIndex = shuffle[nTest+1:lengthData]
testDataSet = cars[shuffle[(nTrain+1):lengthData],]
normalizedTestDataSet = carsNormalized[shuffle[(nTrain+1):lengthData],]

head(trainDataSet)

head(normalizedTrainDataSet)

```

# what factors may be in play??
  
```{r}
linearModel = glm(sales~ age + gender + miles + debt + income, data = trainDataSet)
summary(linearModel)
print(linearModel)
plot(linearModel)
```

 leave gender out  
 
# Traditional stat approach: Logistic Regression  

```{r}
logisticModel = glm(salesTgt ~ age + miles + debt + income, family=binomial, data = trainDataSet)
summary(logisticModel)
plot(logisticModel)
b = as.numeric(logisticModel$coefficients)

```

## quick look- general probability of a sale  
print probabilities of sale based on age from logistic regression

```{r}

averageDebt = mean(testDataSet$debt)
averageIncome = mean(testDataSet$income)
averageMiles = mean(testDataSet$miles)
xx = b[1]+b[2]*testDataSet$age + b[3]*averageMiles +
  b[4]*averageDebt + b[5]*averageIncome
pred.prob = 1./(1+exp(xx))

plot(testDataSet$age, pred.prob, ylab='prob of sale',
     xlab='age')
```

## Cross tabulation of logistic regression predictions  
columns are 'truth,' rows are predictions

```{r echo=FALSE}
# set up the table column
xTabLogis = matrix(0,nrow=2, ncol=2)
probSale = 0.2
for(i in 1:nTest){
  
  if((pred.prob[i]>probSale) && (testDataSet$salesTgt[i]==1) )          xTabLogis[2,2] = xTabLogis[2,2]+1
     
  if((pred.prob[i]>probSale) && (testDataSet$salesTgt[i]==0))
     xTabLogis[2,1] = xTabLogis[2,1]+1

  if((pred.prob[i]<=probSale) && (testDataSet$salesTgt[i]==0))
     xTabLogis[1,1] = xTabLogis[1,1]+1

  if((pred.prob[i]<=probSale) && (testDataSet$salesTgt[i]==1))
     xTabLogis[1,2] = xTabLogis[1,2]+1
      
}

print(xTabLogis)
print('columns are truth in data set, ')
print('rows are predicted by logistic regression ')

accuracyLogis = (xTabLogis[1,1]+xTabLogis[2,2])/nTest
print(paste('Overall accuracy: ', round(accuracyLogis,2)))

```


# On with the NN  

## Use the normalized data!

```{r SalesNet}
# set up the model using the training set
salesNet = neuralnet(salesTgt~age+miles+debt+income, data = normalizedTrainDataSet, hidden=c(2), stepmax=1.e5, rep=5)

```

plot..

```{r  fig.keep='all', fig.width=100}
plot(salesNet)
```

## cross tabulation for the NN

NN cross tabulation; columns are 'truth', rows are predictions

```{r echo=FALSE}
xTabNN = matrix(0,nrow=2, ncol=2)

for(i in 1:nTest){
  
  if((salesNet$response[i]==1) && (testDataSet$salesTgt[i]==1) )          xTabNN[2,2] = xTabNN[2,2]+1
     
  if((salesNet$response[i]==1) && (testDataSet$salesTgt[i]==0))
     xTabNN[2,1] = xTabNN[2,1]+1

  if((salesNet$response[i]==0) && (testDataSet$salesTgt[i]==0))
     xTabNN[1,1] = xTabNN[1,1]+1

  if((salesNet$response[i]==0) && (testDataSet$salesTgt[i]==1))
     xTabNN[1,2] = xTabNN[1,2]+1
      
}

print(xTabNN)
print('columns are truth in data set, ')
print('rows are predicted by logistic regression ')

accuracyNN = (xTabNN[1,1]+xTabNN[2,2])/nTest
print(paste('Overall accuracy: ', round(accuracyNN,3)))


```

## look at a few of the fit values..

```{r}
randomTest = sample(1:length(normalizedTestDataSet$salesTgt), 15)

x = data.frame(age=normalizedTestDataSet$age[randomTest],     miles=normalizedTestDataSet$miles[randomTest],
    debt= normalizedTestDataSet$debt[randomTest], 
    income= normalizedTestDataSet$income[randomTest])

print(x)

print(paste('actual = ', normalizedTestDataSet$salesTgt[randomTest], '  predicted: ', salesNet$response[randomTest]))


```


